Mining Interaction Log Data in a Creative Arts Practice using Transition Matrices

### 2015-09-24 flux-entropy-paper/flux-entropy-paper.tex
adjusted a few figure captions

\begin{abstract}

  In this note we present an analysis of an interaction data archive
  associated with a long-term practice-led research program
  investigating ensemble musical performance practice on touch
  screens. We use transition matrix measures to distinguish between
  sessions of different contexts, and the internal structure of these
  sessions. The data analysis proceeds by first classifying these data
  into higher level gesture types and summarising the performers'
  transitions between these gestures. We find that two
  transition-matrix measures, flux, and entropy, can be used to
  distinguish different performance types and contexts, as well as the
  internal structure of the performances in our corpus.
\end{abstract}

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figures/three-performance-contexts}
  \caption{Collaborative musical interaction with our touch-screen
    interfaces in two different {\em contexts} and two different {\em instrumentations}: (left to right) rehearsal
    in a studio with iPads only,
    concert performance with iPads and percussion instruments, and concert performance with 
    iPads only.\label{fig:three-performance-contexts}}
\end{figure*}

@@

\begin{table}
\centering
\begin{tabular}{l|llll}
\hline
            & Total & Min  & Median   & Max     \\ 
\hline
N           & 95    &      &          &         \\
%Record size & 306MB     &         &        &   \\
Length & 13H9M19S & 1M19S & 7M26S & 22M20S \\
Participants& 353   & 2    & 4        & 9        \\
Flux &   & 0.02239 & 0.1445 & 0.3246\\
Entropy &   & 0.7683 & 3.342 & 4.442\\          
\hline
\end{tabular}
\caption{
  Representative statistics from the corpus of musical touch-screen
  interaction sessions described in this note.  Each record in the corpus
  consists of low level touch-data (such as touches begin, touches move, touches end) recorded in CSV format. 
  Many of the participants were
  present in multiple sessions, but this data was not tracked in
  our (deidentified) corpus.\label{corpus-table}}
\end{table}

@@

\begin{figure} \centering
  \includegraphics[width=\linewidth]{figures/context-flux-entropy-boxplot}
  \caption{Distributions of flux and entropy values by performance
context. Study and rehearsal sessions had the highest flux for
improvisations. Compositions had lower
flux in rehearsals than concerts.
\label{fig:flux-entropy-boxplot}}
\end{figure}

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
          \includegraphics[width=\linewidth]{figures/type-section-flux-entropy}
          \caption{Sessions divided by session type. \label{fig:type-section-flux-entropy}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
          \includegraphics[width=\linewidth]{figures/context-section-flux-entropy}
          \caption{Sessions divided by session context. \label{fig:context-section-flux-entropy}}
    \end{subfigure}
    \caption{The flux and entropy of different interaction sessions
      divided into beginning, middle, and end.}
\end{figure*}

>>>>

\begin{abstract}
  In this note we present an analysis of an interaction data archive
  associated with a long-term practice-led research program
  investigating ensemble musical performance practice on touch
  screens. We use transition matrix measures to distinguish between
  sessions of different contexts, and the internal structure of these
  sessions. The data analysis proceeds by first classifying these data
  into higher level gesture types and summarising the performers'
  transitions between these gestures. We find that two
  transition-matrix measures, flux, and entropy, can be used to
  distinguish different performance types and contexts, as well as the
  internal structure of the performances in our corpus.
\end{abstract}

@@

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figures/three-performance-contexts}
  \caption{Collaborative musical interaction with our touch-screen
    interfaces in two different {\em contexts} and two different {\em
      instrumentations}: (left to right) a rehearsal with iPads only,
    concert performance with iPads and percussion instruments, and
    concert performance with iPads
    only.\label{fig:three-performance-contexts}}
\end{figure*}

@@

\begin{table}
\centering
\begin{tabular}{l|llll}
\hline
            & Total & Min  & Median   & Max     \\ 
\hline
N           & 95    &      &          &         \\
%Record size & 306MB     &         &        &   \\
Length & 13H9M19S & 1M19S & 7M26S & 22M20S \\
Participants& 353   & 2    & 4        & 9        \\
Flux &   & 0.02239 & 0.1445 & 0.3246\\
Entropy &   & 0.7683 & 3.342 & 4.442\\          
\hline
\end{tabular}
\caption{
  Representative statistics from the corpus of musical touch-screen
  interaction sessions described in this note.  Each record in the corpus
  consists of low level touch-data (such as touches begin, touches move, touches end) recorded in CSV format. 
  Many of the participants were
  present in multiple sessions, but this data was not tracked.\label{corpus-table}}
\end{table}

@@

\begin{figure} \centering
  \includegraphics[width=\linewidth]{figures/context-flux-entropy-boxplot}
  \caption{Distributions of flux and entropy values by performance
context. Recording and rehearsal sessions had the highest flux for
improvisations. Compositions had lower
flux in rehearsals than concerts.
\label{fig:flux-entropy-boxplot}}
\end{figure}

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
          \includegraphics[width=\linewidth]{figures/type-section-flux-entropy}
          \caption{Sessions divided by performance type. \label{fig:type-section-flux-entropy}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\linewidth}
        \centering
          \includegraphics[width=\linewidth]{figures/context-section-flux-entropy}
          \caption{Improvisation performances divided by session context. \label{fig:context-section-flux-entropy}}
    \end{subfigure}
    \caption{The flux and entropy of different interaction sessions
      divided into beginning, middle, and end.}
\end{figure*}
